---
title: "BetaDivPLOTS"
author: "Emily Bean"
date: "6/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

## READ IN AND PREPARE DATA
if(!require(ggordiplots)) {
  require(devtools)
  devtools::install_github("jfq3/ggordiplots")
  require(ggordiplots)
}
require(vegan)
require(tidyverse)
require(phyloseq)
require(pairwiseAdonis)
require(RColorBrewer)
# load phyloseq objects and metadata
if(!require(repmis)) {
  install.packages("repmis")
  require(repmis)
}
source_data("https://github.com/EmilyB17/soils-micro/blob/master/data/phyloseq16S_ITS.RData?raw=true")

set.seed(123)

# set brewer colors
brewer.pal(n = 8, name = "Dark2")
treatmentcols <- c("HDG" = "#1B9E77", "LDG" =  "#D95F02", "NG" = "#7570B3")
treatmentcols2 <- c("HI" = "#1B9E77", "LO" =  "#D95F02", "NO" = "#7570B3")
timecols <- c("PRE" = "#E7298A", "24H" = "#66A61E", "1WK" = "#E6AB02", "4WK" = "#A6761D")

```

# ITS

### Wrangle Data & Ordination

```{r}

# Add EEA C cycling
# also make sure ADONIS variables are factors
sdITS18 <- sdITS18 %>% 
  mutate(Ccycl = AG + BG + BX + CBH,
         Ncycl = NAG + LAP,
         CNrat = Ccycl / Ncycl,
         Plot = as.factor(Plot),
         Block = as.factor(Block))

## conglomerate OTUs by Phyla and add to metadata
# agglomerate data by Phyla
tg <- tax_glom(psITS18, taxrank = rank_names(psITS18)[2], NArm = TRUE)
# name the phyla
taxa_names(tg) <- tg@tax_table@.Data[,2]
# make a dataframe
df <- as.data.frame(otu_table(tg))
# create relative abundance of Phyla for each sample
dc <- decostand(df, method = "total", MARGIN = 1)
# combine with sample data
sdITS18 <- cbind(dc, sdITS18) 
# remove Plot and Block (not interested in differences)
sdITS18$Plot <- NULL
sdITS18$Block <- NULL
# make sure GrazeTime is ordered
sdITS18$GrazeTime <- factor(sdITS18$GrazeTime, ordered = TRUE,
                            levels = c("PRE", "24H", "1WK", "4WK"))

# statistical testing
dis <- distance(psITS18T, method = "bray")
adonis(dis ~ Treatment * GrazeTime,
       data = sdITS18, permutations = 999) # all are significant except Interactions

# get Treatment effects
 pairwise.adonis2(dis ~  Treatment * GrazeTime,
                             data = sdITS18,
                             p.adjust.m = "bon", perm = 1000)
 
 # get Time effects
  pairwise.adonis2(dis ~  GrazeTime * Treatment,
                             data = sdITS18,
                             p.adjust.m = "bon", perm = 1000)

# ordinate
ord <- ordinate(psITS18T, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = sdITS18, perm = 1000, na.rm = TRUE)
eft #BG, PEROX, Ascomycota, Basidiomycota, factors: Plot, Block, Treatment


```

There are significant differences between each Treatment; NO vs LO, NO vs HI, and LO vs HI (this also has a time effect).There is also a Treatment effect in 24H vs 1Wk, 1WK vs 4WK, PRE vs 1WK, and 24H vs 4WK (also has a Time effect). (PRE vs 24H is not significant)

To tease apart these differences, we plot an NMDS and do an adonis test for all 3 treatments by Time.

## One NMDS for all data

```{r}

# ordinate
ord <- ordinate(psITS18T, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = sdITS18, perm = 1000, na.rm = TRUE)
eft #BG, PEROX, Ascomycota, Basidiomycota, factors: Plot, Block, Treatment


## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = sdITS18$Treatment,
         Time = sdITS18$GrazeTime,
         Type = sdITS18$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = sdITS18$Treatment, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## 4. PLOT BY TREATMENT
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, shape = Time, color = Treatment), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Treatment", shape = "Time")
# save
#ggsave("./data/plots/NMDS-fungi-Treatment-labels.png", plot = last_plot(), dpi = "print")

## WITHOUT LABELS - also need to rename factor levels
coords$Treatment <- fct_recode(coords$Treatment, HDG = "HI", LDG = "LO", NG = "NO")
df_ell$Group <-fct_recode(df_ell$Group, HDG = "HI", LDG = "LO", NG = "NO")

## 4. PLOT BY TREATMENT
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, shape = Time, color = Treatment), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  #geom_segment(data = sigspecies,
   #            aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
    #           arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
 # geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
  #          size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Treatment", shape = "Time") +
  # colors for each treatment
  scale_color_manual(values = treatmentcols)
# save
#ggsave("./data/plots/NMDS-fungi-Treatment-nolabels.png", plot = last_plot(), dpi = "print")

```


## Individual plots

#### HDG

```{r}

## HDG

tax <- subset_samples(psITS18T, Treatment == "HI")
dat <- dat %>% filter(Treatment == "HI")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## no differences

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# Plot - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## Plot by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## Plot by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
     #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-fungi-HDG-nolabels.png", plot = last_plot(), dpi = "print")

```

#### LDG

```{r}




## Ordinate and plot each individual treatment


## LDG

tax <- subset_samples(psITS18T, Treatment == "LO")
dat <- sdITS18 %>% filter(Treatment == "LO")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## PRE vs 24H, 24H vs 4WK

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# Plot - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## Plot by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## Plot by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
     #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-fungi-LDG-nolabels.png", plot = last_plot(), dpi = "print")

```

#### NG

```{r}




## Ordinate and plot each individual treatment


## NG

tax <- subset_samples(psITS18T, Treatment == "NO")
dat <- sdITS18 %>% filter(Treatment == "NO")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## no differences

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# PNOt - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## PNOt by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## PNOt by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
     #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-fungi-NG-nolabels.png", plot = last_plot(), dpi = "print")

```

# 16S

### Wrangle data and ordination

```{r}

# Add EEA C cycling
# also make sure ADONIS variables are factors
sd16S18 <- sd16S18 %>% 
  mutate(Ccycl = AG + BG + BX + CBH,
         Ncycl = NAG + LAP,
         CNrat = Ccycl / Ncycl,
         Plot = as.factor(Plot),
         Block = as.factor(Block))

## conglomerate OTUs by Phyla and add to metadata
# agglomerate data by Phyla
tg <- tax_glom(ps16S18, taxrank = rank_names(ps16S18)[2], NArm = TRUE)
# name the phyla
taxa_names(tg) <- tg@tax_table@.Data[,2]
# make a dataframe
df <- as.data.frame(otu_table(tg))
# create relative abundance of Phyla for each sample
dc <- decostand(df, method = "total", MARGIN = 1)
# combine with sample data
sd16S18 <- cbind(dc, sd16S18) 
# remove Plot and Block (not interested in differences)
sd16S18$Plot <- NULL
sd16S18$Block <- NULL
# make sure GrazeTime is ordered for plots
sd16S18$GrazeTime <- factor(sd16S18$GrazeTime, ordered = TRUE,
                            levels = c("PRE", "24H", "1WK", "4WK"))

# statistical testing
dis <- distance(ps16S18T, method = "bray")

adonis(dis ~ Treatment * GrazeTime,
       data = sd16S18, permutations = 999) # only GrazeTime
pairwise.adonis2(dis ~  GrazeTime * Treatment,
                 data = sd16S18,
                 p.adjust.m = "bon", perm = 1000) # all GrazeTime comparisons are significant
# ordinate
ord <- ordinate(ps16S18T, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = sd16S18, perm = 1000, na.rm = TRUE)
eft #BG, PEROX, Ascomycota, Basidiomycota, factors: Plot, Block, Treatment

```

#### HDG

```{r}


## HDG

tax <- subset_samples(ps16S18T, Treatment == "HI")
dat <- sd16S18 %>% filter(Treatment == "HI")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## PRE vs 24H, 24H vs 4WK

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# Plot - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## Plot by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## Plot by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
  #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-bacteria-HDG-nolabels.png", plot = last_plot(), dpi = "print")


```


#### LDG

```{r}


## LDG

tax <- subset_samples(ps16S18T, Treatment == "LO")
dat <- sd16S18 %>% filter(Treatment == "LO")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## PRE vs 24H, 24H vs 4WK, PRE vs 1WK

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# Plot - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## Plot by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## Plot by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
  #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-bacteria-LDG-nolabels.png", plot = last_plot(), dpi = "print")


```

#### NG

```{r}


## NG

tax <- subset_samples(ps16S18T, Treatment == "NO")
dat <- sd16S18 %>% filter(Treatment == "NO")

dis <- distance(tax, method = "bray")

pairwise.adonis2(dis ~  GrazeTime,
                 data = dat,
                 p.adjust.m = "bon", perm = 1000) ## NONE

# Ordinate
ord <- ordinate(tax, method = "NMDS", distance = "bray", k = 2, 
                trymax = 1000, previousBest = TRUE) 
stressplot(ord, dis) #OK with 2 dimensions
eft <- envfit(ord, env = dat, perm = 1000, na.rm = TRUE)
eft

# Plot - WITH LABELS
## 1.  get NMDS coordinates from ordination object
coords <- as.data.frame(scores(ord, display = "sites")) %>% 
  # and add grouping variable
  mutate(Treatment = dat$Treatment,
         Time = dat$GrazeTime,
         Type = dat$sample_type)

## 2.  get significant envfit objects to plot as arrows
spp <- as.data.frame(scores(eft, display = "vectors")) 
spp$species <- rownames(spp)
# subset to only p < 0.05 vectors to de-clutter 
sigs <- data.frame(eft$vectors$arrows[eft$vectors$pvals < 0.05,])
sigspecies <- rownames(sigs)
# subset
sigspecies <- spp %>% filter(species %in% sigspecies)

## 3. Pull ellipse information using gg_ordiplot package
# save gg_ordiplot object to get ellipse values
plot <-  gg_ordiplot(ord, groups = dat$GrazeTime, label = FALSE, plot = FALSE)
# get ellipse coordinates
df_ell <- plot$df_ellipse
# get label coordinates for ellipse centers
NMDS.mean <- plot$df_mean.ord
# pull NMDS coordinates
ord.data <- plot$df_ord 

## Plot by TIME - WITH LABELS
## create in ggplot2
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
            size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)

## Plot by Time - WITHOUT LABELS
ggplot(data = coords, aes(x = NMDS1, y = NMDS2)) + # label axises automatically
  # ELLIPSES
  geom_path(data = df_ell, aes(x = x, y = y, color = Group), show.legend = FALSE) +
  # ORDINATION POINTS
  geom_point(data = coords, aes(x = NMDS1, y = NMDS2, color = Time), size = 1) +
  # GROUP NAMES AT ELLIPSE CENTER 
  #annotate("text",x = NMDS.mean$x, y = NMDS.mean$y,label=NMDS.mean$Group, size = 5) +
  # ENVFIT ARROWS
  geom_segment(data = sigspecies,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  # ARROW TEXT
  #geom_text(data = sigspecies, aes(x = NMDS1, y = NMDS2, label = species),
  #       size = 3) +
  # SCALE CORRECTLY
  coord_fixed() +
  # THEME
  theme_bw() +
  # CHANGE LEGEND TITLE
  labs(color = "Sampling Time") +
  # add constant colors for each Time
  scale_color_manual(values = timecols)
# save
#ggsave("./data/plots/NMDS-bacteria-NG-nolabels.png", plot = last_plot(), dpi = "print")

```

